---
title: "Interpolation Test"
author: "Roque Arcudia Hernandez"
date: "August 14, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Setting up the date to analyze:

```{r date_setup}
source('WeatherTest.R')
# Global var for table display:
headSize<-10
  
myDate <- as.Date("2016-12-31")
myDate
```

Doing a query on the weather data for that date as a Zoo (printing as data frame or we don't get the time):

```{r zoo_query}
myZoo<- getPWSZoo("IYUCATNT2",myDate)
knitr::kable(head(as.data.frame(myZoo),headSize))
```

Defining the resampling as a range between the start and end time separated by 5 min. This is close to what the dataset captures.

```{r resampling}
resample <- seq(start(myZoo),end(myZoo),by="5 mins")
# These 2 lines resample each 5 mins of the day:
#myDay <- floor_date(start(myZoo),unit="day")
#resample <- seq(myDay,by="5 mins", len = (24*60/5))
resample
str(start(myZoo))
```

Creating an empty zoo with the previous resampling sequence that will be used for the merge:

```{r emptyzoo}
emptyZoo <- zoo(,resample)
```

Creating a new merged Zoo where elements from the original Zoo are present as well as elements of an empty Zoo created with the resample sequence. The new elementes will have NAs

```{r newsamples}
mergedZoo <- merge(myZoo,emptyZoo,all=TRUE)
knitr::kable(head(as.data.frame(mergedZoo),headSize))
```

Interpolate the NAs with other values:

```{r interpolate}
mergedZoo <- na.approx(mergedZoo)
knitr::kable(head(as.data.frame(mergedZoo),headSize))
```

Merge again with empty Zoo to keep only the times that fit into the 5 min interval in case it is very irregular sampling in the original data set.

```{r remerge}
mergedZoo <- merge(mergedZoo,emptyZoo,all=FALSE)
knitr::kable(head(as.data.frame(mergedZoo),headSize))
```

This is the plot of the original zoo:

```{r plot_original}
plot(myZoo$WindSpeedMPH,type="b")
```

Interpolated:

```{r plot_inter}
plot(mergedZoo$WindSpeedMPH,type="b")
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
