---
title: "Assignment 3"
author: "Roque Arcudia Hernandez"
date: "July 21, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this homework, you will do some data analysis using R for the Forest Fire Data described http://archive.ics.uci.edu/ml/machine-learning-databases/forest-fires/forestfires.csv

The dataset is used to find the relationship between the burned area of forest fires and meteorological data.

**1. Import the data into R.**
```{r import_data}
forestfires <- read.csv('http://archive.ics.uci.edu/ml/machine-learning-databases/forest-fires/forestfires.csv')
```

**a. How many observations are there in the
dataset?**
```{r size_dataset}
nrow(forestfires)
```

**b. How many observations are there with a fire
(i.e., area>0)**
```{r size_with_fire}
sum(forestfires$area>0)
```

**c. How many observations are there with a rain
(i.e., rain>0)**
```{r size_with_rain}
sum(forestfires$rain>0)
```

**d. How many observations are there with both a
fire and a rain?**
```{r size_with_fire_rain}
sum(forestfires$area>0 & forestfires$rain>0)
```

**2. Show the columns month, day, area of the all the observations.**
```{r month_day_area}
knitr::kable(subset(forestfires,select=c('month', 'day', 'area')))
```

**3. Show the columns month, day, area of the observations with a fire**
```{r month_day_area_fire}
knitr::kable(subset(forestfires,area>0,select=c('month', 'day', 'area')))
```

**4. How large are the five largest fires (i.e., having largest area)**
```{r five_largest}
largest_fires <- head(sort(forestfires$area,TRUE),5)
largest_fires
```

**a. What are the corresponding month, temp, RH, wind, rain area?**
```{r five_largest_subset}
knitr::kable(subset(forestfires,area %in% largest_fires,select=c('month', 'temp', 'RH', 'wind', 'rain', 'area')))
```

**b. Reorder factor levels of month to be from Jan to Dec. Add one column to the data indicating whether a fire occurred for each observation (True for area >0 and False for area ==0)**

before:
```{r before}
levels(forestfires$month)
```

Reordering month factor and adding a new column fire:

```{r transformations}
forestfires<-transform(forestfires,
  month=factor(month,levels=c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec"),ordered=TRUE),
  fire=area>0)
```
Verification:
```{r verification}
levels(forestfires$month)
knitr::kable(forestfires[order(forestfires$month),])
```
