---
title: "Assignment 1"
author: "Roque Arcudia Hernandez"
date: "July 10, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=4, fig.height=3, fig.align='center')
```

**1. What are the measures of central tendency of data? What is the order of these measures for skewed data?**

*Mean* : is the sum of all the values in a data set divided by the number of elements in the set. It uses the symbol $\mu$ for a population and $\bar{x}$ for a sample mean. On a data set $X$ of size $N$:
$$\frac{1}{N}\sum_{i=1}^N X_i$$

*Median* : Is the middle value of a data set when ordered. If the data set has odd number of elements it is simply the middle element. If the data set has even number of elements then it is the average of values in the positions $\frac{n}{2}$ and $\frac{n}{2}+1$ of the ordered dataset, where $n$ is the number of elements.

*Mode* : Is the most common element in a data set.

*Skewed Right* : Also called positive skew, the right tail will be longer.

```{r skewed_right, echo=FALSE}
hist(rbeta(100000,1.5,5),xlab="",ylab="",axes=FALSE,main="")
```

The typical order of the measures of central tendency for skewed right distributions is:
$$
Mode < Median < Mean
$$
*Skewed left* : Also called negative skew, the left tail will be longer.

```{r skewed_left, echo=FALSE}
hist(rbeta(100000,5,1.5),xlab="",ylab="",axes=FALSE,main="")
```

The typical order of the measures of central tendency for skewed left distributions is:
$$
Mean < Median < Mode
$$

**2. How can you measure variation of data?**

With the following:

*Range* : Difference between the maximum and minimum value in a distribution.

*Inter Quartile Range* : Difference between the 3rd quartile and the 1st quartile.

*Variance* : is a measure of how disperse is the data set with respect to the mean value.

From [stattrek.com](http://stattrek.com/): variance is the average squared deviation from the population mean, as defined by the following formula:
$$\sigma^2 = \frac{1}{N}\sum_{i=1}^N (X_i - \mu)^2$$

In case of a sample it gets calculated as:
$$s^2 = \frac{1}{n-1}\sum_{i=1}^n (x_i - \bar{x})^2$$


*Standard Deviation* : Is the square root of the variance. For a population:
$$\sigma = \sqrt{\frac{\sum_{i=1}^N (X_i - \mu)^2}{N}}$$

And for a sample:
$$s = \sqrt{\frac{\sum_{i=1}^n (x_i - \bar{x})^2}{n-1}}$$

**3. Create a vector of positive odd integers less
than 100**
```{r odd_integers}
odd <- seq(1,99,by=2)
odd
```
**4. Remove the values greater than 60 and less
than 80**
```{r no_greater_than_60_and_less_than_80}
odd <- odd[!( odd>60 & odd<80)]
odd
```

**5. Find the five number summary of the
remaining set of values**
```{r five_num_summary}
fnum <- fivenum(odd)
names(fnum)<-c("Min.","1st Qu.","Median", "3rd Qu.","Max.")
fnum
```

**6. Consider the following vector of values.
X <-{8, 14, 9, 15, NA, 8,13, 2,9, NA}
Write an R program to return the positions of the missing values**

```{r NA_pos}
X <- c(8, 14, 9, 15, NA, 8,13, 2,9, NA)
NAPos <- which(is.na(X))
NAPos
```

**7. Write an R program to count the number of non-missing values**
```{r count_non_NA}
nonNACount <- sum(!is.na(X))
nonNACount
```

**8. Write an R program to replace the missing values with the mean of the non-missing values.**
_Note : Reusing NAPos from 6._
```{r replace_NA}
X[NAPos] <- mean(X[!is.na(X)])
X
```

**9. Load mtcars data. Write an R program that will rearrange the rows of the data frame so that they are sorted by the value of Horsepower.**
```{r mtcars_order_by_hp}
mtcarsByHP <- mtcars[order(mtcars$hp),]
mtcarsByHP
```

**10. Write an R program to count the number of observations with cylinders greater than 4 and gear greater than 3.**

```{r obs_filter}
cylGt4GearGt3 <- sum( mtcars$cyl>4 & mtcars$gear>3 )
cylGt4GearGt3
```
