---
title: "Assignment 1"
author: "Roque Arcudia"
date: "July 10, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## ASSIGNMENT # 1
1. What are the measures of central tendency of data? What is the order of these measures for skewed data?
2. How can you measure variation of data?
3. Create a vector of positive odd integers less
than 100
```{r odd_integers}
odd <- seq(1,99,by=2)
odd
```
4. Remove the values greater than 60 and less
than 80
```{r no_greater_than_60_and_less_than_80}
odd <- odd[!( odd>60 & odd<80)]
odd
```

5. Find the five number summary of the
remaining set of values
```{r five_num_summary}
fnum <- fivenum(odd)
names(fnum)<-c("Min.","1st Qu.","Median", "3rd Qu.","Max.")
fnum
```

6. Consider the following vector of values.
X <-{8, 14, 9, 15, NA, 8,13, 2,9, NA}
Write an R program to return the positions of the missing values

```{r NA_pos}
X <- c(8, 14, 9, 15, NA, 8,13, 2,9, NA)
NAPos <- which(is.na(X))
NAPos
```

7. Write an R program to count the number of non-missing values
```{r count_non_NA}
nonNACount <- sum(!is.na(X))
nonNACount
```

8. Write an R program to replace the missing values with the mean of the non-missing values.
```{r replace_NA}
X[NAPos] <- mean(X[!is.na(X)])
X
```

9. Load mtcars data. Write an R program that will rearrange the rows of the data frame so that they are sorted by the value of Horsepower.
```{r mtcars_order_by_hp}
mtcarsByHP <- mtcars[order(mtcars$hp),]
mtcarsByHP
```

10. Write an R program to count the number of observations with cylinders greater than 4 and gear greater than 3.

```{r obs_filter}
cylGt4GearGt3 <- sum( mtcars$cyl>4 & mtcars$gear>3 )
cylGt4GearGt3
```
